Simulate regression example

### 2022-10-08 notebook/_posts/2021-08-03-simulate-regression-example/simulate-regression-example.Rmd
Revise figure of regression simulation

species <- str_c("Species ", seq(n_species))
```

@@

```{r}
p_species_all <- abun %>%
  ggplot(aes(x, log2_abundance, color = type)) +
  labs(y = "Log abundance ") +
  facet_grid(.otu~type, scales = "fixed") +
  geom_quasirandom(alpha = 0.3, groupOnX = TRUE) +
  # stat_summary(fun.data = "mean_cl_boot", geom = "pointrange") +
  stat_summary(geom = "point") +
  stat_smooth(aes(x = as.integer(x)),
    method = "lm", size = 0.9, fill = 'grey', se = FALSE
  ) +
  theme(legend.position = "none")
p_species_all 
```

```{r}
p_mean_eff <- mean_eff %>%
  ggplot(aes(x, log2(mean_efficiency))) +
  labs(y = "Log efficiency ") +
  geom_quasirandom(alpha = 0.3, groupOnX = TRUE) +
  # stat_summary(fun.data = "mean_cl_boot", geom = "pointrange") +
  stat_summary(geom = "point") +
  stat_smooth(aes(x = as.integer(x)),
    method = "lm", size = 0.9, color = 'black', fill = 'grey', se = FALSE
  )
p_mean_eff 
```

@@

```{r, fig.dim = c(5.5, 7)}
# params for arrows showing error
delta <- lm_results_mean_eff %>% filter(term == "x1") %>% pull(estimate)
start <- lm_results_slope %>% filter(type == 'Actual', .otu == 'Species 9') %>%
  pull(estimate)

p_coef_ci <- lm_results_slope %>% 
  ggplot(aes(y = .otu, x = estimate, color = type)) +
  labs(x = "Log fold difference", y = NULL) +
  geom_vline(xintercept = 0, color = "grey") +
  geom_pointinterval(aes(xmin = conf.low, xmax = conf.high)) +
  theme(legend.position = 'top') +
  guides(color = guide_legend(reverse = TRUE)) +
  annotate(
    geom = 'segment', color = "darkgrey",
    arrow = grid::arrow(length = unit(0.1, "inches")), 
    x = start, xend = start - delta, 
    y = 10.5, yend = 10.5
  ) +
  coord_cartesian(clip = 'off')

@@

p_coef_ci_mean_eff <- lm_results_mean_eff %>% 
  filter(term == "x1") %>%
  ggplot(aes(y = "Mean efficiency", x = estimate)) +
  expand_limits(
    x = c(
      min(lm_results_slope$conf.low), 
      max(lm_results_slope$conf.high)
  )) +
  labs(x = "Log fold difference", y = NULL) +
  geom_vline(xintercept = 0, color = "grey") +
  geom_point() +
  annotate(
    geom = 'segment', color = "darkgrey",
    arrow = grid::arrow(length = unit(0.1, "inches")), 
    x = 0, xend = delta, 
    y = 1.0, yend = 1.0
  )

p_coef_dot <- lm_results_slope %>% 
  ggplot(aes(y = type, x = estimate, fill = type)) +
  labs(x = "Log fold difference", y = "Type") +
  geom_vline(xintercept = 0, color = "grey") +
  stat_dots()

@@

```{r}
species_to_plot <- str_c("Species ", c('9', '10', '5', '7', '2'))
p_species_focal <- p_species_all
p_species_focal$data <- p_species_focal$data %>% 
  filter(.otu %in% species_to_plot)

@@

```{r main-figure, fig.dim = c(7, 7)}
(p_mean_eff1 + ggtitle("Mean efficiency")) +
  (p_coef_ci_mean_eff + ggtitle("LFD in mean efficiency")) + 
  (p_species_focal + ggtitle("Actual and measured\nabundances of select species")) + 
  (p_coef_ci1 + ggtitle("Estimated LFDs of all species") +
    theme(legend.box.margin = margin(b = -15))
  ) + 
  plot_layout(ncol = 2, heights = c(0.2, 1)) +
  plot_annotation(tag_levels = 'A') &
  colorblindr::scale_color_OkabeIto() &
  colorblindr::scale_fill_OkabeIto() &
  theme(
    plot.title = element_text(face = "plain")
  )
```

@@

>>>>

species <- str_c("Sp. ", seq(n_species))
```

@@

```{r}
p_species_all <- abun %>%
  ggplot(aes(x, log2_abundance, color = type)) +
  labs(y = "Log efficiency", x = "Condition (x)") +
  facet_grid(.otu~type, scales = "fixed") +
  geom_quasirandom(alpha = 0.3, groupOnX = TRUE) +
  # stat_summary(fun.data = "mean_cl_boot", geom = "pointrange") +
  stat_summary(geom = "point", fun.data = mean_se) +
  stat_smooth(aes(x = as.integer(x)), formula = y~x,
    method = "lm", size = 0.9, fill = 'grey', se = FALSE
  ) +
  theme(legend.position = "none")
p_species_all 
```

```{r}
p_mean_eff <- mean_eff %>%
  ggplot(aes(x, log2(mean_efficiency))) +
  labs(y = "Log efficiency", x = "Condition (x)") +
  geom_quasirandom(alpha = 0.3, groupOnX = TRUE) +
  # stat_summary(fun.data = "mean_cl_boot", geom = "pointrange") +
  stat_summary(geom = "point", fun.data = mean_se) +
  stat_smooth(aes(x = as.integer(x)), formula = y~x,
    method = "lm", size = 0.9, color = 'black', fill = 'grey', se = FALSE
  )
p_mean_eff 
```

@@

```{r, fig.dim = c(5.5, 7)}
# params for arrows showing error
delta <- lm_results_mean_eff %>% filter(term == "x1") %>% pull(estimate)
start <- lm_results_slope %>% filter(type == 'Actual', .otu == 'Sp. 9') %>%
  pull(estimate)

p_coef_ci <- lm_results_slope %>% 
  ggplot(aes(y = .otu, x = estimate, color = type)) +
  labs(x = "Mean LFD", y = NULL, color = "Type") +
  geom_vline(xintercept = 0, color = "grey") +
  geom_pointinterval(aes(xmin = conf.low, xmax = conf.high)) +
  theme(legend.position = 'top') +
  guides(color = guide_legend(reverse = TRUE)) +
  annotate(
    geom = 'segment', color = "darkred",
    arrow = grid::arrow(length = unit(0.1, "inches")), 
    x = start, xend = start - delta, 
    y = 10.5, yend = 10.5,
    size = .7
  ) +
  annotate(
    geom = 'text', color = "darkred",
    label = 'Effect of bias',
    x = start - delta/2,
    y = 10.8,
  ) +
  coord_cartesian(clip = 'off')

@@

p_coef_ci_mean_eff <- lm_results_mean_eff %>% 
  filter(term == "x1") %>%
  ggplot(aes(y = "Mean efficiency", x = estimate)) +
  expand_limits(
    x = c(
      min(lm_results_slope$conf.low), 
      max(lm_results_slope$conf.high)
  )) +
  labs(x = "Mean LFD", y = NULL) +
  geom_vline(xintercept = 0, color = "grey") +
  geom_point() +
  annotate(
    geom = 'segment', color = "darkred",
    arrow = grid::arrow(length = unit(0.1, "inches")), 
    x = 0, xend = delta, 
    y = 1.0, yend = 1.0,
    size = .7
  )

p_coef_dot <- lm_results_slope %>% 
  ggplot(aes(y = type, x = estimate, fill = type)) +
  labs(x = "Mean LFD", y = "Type") +
  geom_vline(xintercept = 0, color = "grey") +
  stat_dots()

@@

```{r}
species_to_plot <- str_c("Sp. ", c('9', '10', '5', '7', '2'))
p_species_focal <- p_species_all
p_species_focal$data <- p_species_focal$data %>% 
  filter(.otu %in% species_to_plot)

@@

```{r main-figure, fig.dim = c(7, 7)}
(p_mean_eff1 + ggtitle("Mean efficiency")) +
  (p_coef_ci_mean_eff + ggtitle("LFD in mean efficiency")) + 
  (p_species_focal + ggtitle("Actual and measured\nabundances of select species")) + 
  (p_coef_ci1 + ggtitle("Mean LFD estimated from\nactual or measured abundances") +
    theme(legend.box.margin = margin(b = -15))
  ) + 
  plot_layout(ncol = 2, heights = c(0.2, 1)) +
  plot_annotation(tag_levels = 'A') &
  colorblindr::scale_color_OkabeIto() &
  colorblindr::scale_fill_OkabeIto() &
  theme(
    plot.title = element_text(face = "plain")
  )
```

@@

## 2022-10-08 Modified main figure

New panel showing simulated efficiencies and mean LFDs, to replace the panel showing the change in mean efficiency (former panel B).

```{r}
p_params <- species_params %>%
  ggplot(aes(x1, log2_efficiency)) +
  theme(axis.line = element_blank()) +
  # panel_border(remove = TRUE) +
  # theme_minimal_grid() +
  coord_cartesian(clip = 'off') +
  geom_vline(xintercept = 0, color = 'grey') +
  geom_hline(yintercept = 0, color = 'grey') +
  geom_text(
    aes(label = str_c('Sp. ', str_extract(.otu, '[0-9]+'))), 
    size = 4) +
  labs(
    x = 'Expected LFD',
    y = 'Log efficiency'
  )
p_params 
```

Annotate mean efficiency plot,

```{r}
# params for arrow showing change
arrow_params <- mean_eff %>%
  with_groups(x, summarize, across(log2_mean_efficiency, mean)) %>%
  pull(log2_mean_efficiency)

p_mean_eff2 <- p_mean_eff +
  annotate(
    geom = 'segment', color = "darkred",
    arrow = grid::arrow(length = unit(0.1, "inches")), 
    x = 2.5, xend = 2.5, 
    y = arrow_params[1], 
    yend = arrow_params[2], 
    size = .7
  )
p_mean_eff2 
```

Assemble with improved titles,

```{r main-figure-2, fig.dim = c(8, 9)}

(p_params + ggtitle("Simulated log efficiencies\nand expected LFDs")) +
  (p_mean_eff2 + ggtitle("Mean efficiency across\nsamples in each condition")) +
  # (p_coef_ci_mean_eff + ggtitle("LFD in mean efficiency")) + 
  (p_species_focal + ggtitle("Actual and measured\nabundances of select species")) + 
  (p_coef_ci1 + ggtitle("Mean LFD estimated from\nactual or measured abundances") +
    theme(legend.box.margin = margin(b = -15))
  ) + 
  plot_layout(ncol = 2, heights = c(0.4, 1)) +
  plot_annotation(tag_levels = 'A') &
  colorblindr::scale_color_OkabeIto() &
  colorblindr::scale_fill_OkabeIto() &
  theme(
    plot.title = element_text(face = "plain")
  )
```
